---
title: "iFood Random Forest Analysis"
format: docx
editor: visual
---

Our previous client, iFood, believes that a customer's willingness to respond to marketing promotions is a proxy for their overall value as a customer. They have asked for help identifying consumers that are most likely to respond to promotions. More specifically, we are going to perform a *response classification* analysis, where we assign consumers into one of three groups based on their previous campaign response levels - "Highly Responsive", "Occasionally Responsive", and "Non-responsive". Our task is to build a model that predicts which customers fall into each segment.

## Preliminaries and Data Cleaning

Load libraries.

```{r}
library(tidyverse)
library(tidymodels)
```

Import the iFood data.

```{r}
(ifood <- read_csv("ifood.csv"))
```

Drop any observations with missing values.

```{r}
ifood <- ifood |>
  drop_na()
```

Define our dependent variable.

```{r}
ifood <- ifood |>
  mutate(response_total = AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 +
           AcceptedCmp4 + AcceptedCmp5 + Response)
```

Visualize the distribution of how many times consumers have responded to previous marketing campaigns.

```{r}
ifood |>
  ggplot(aes(x = response_total)) +
  geom_histogram() +
  labs(x = "Number of Campaign Responses",
       y = "Count",
       title = "Consumer-Level Campaign Response Distribution")
```

Define segment based on above distribution. Convert `segment` into a factor for our model.

```{r}
ifood <- ifood |>
  mutate(
    segment = case_when(
      response_total == 0 ~ "Nonresponsive",
      response_total %in% c(2, 3, 4, 5) ~ "Highly Responsive",
      response_total == 1 ~ "Semi-Responsive"),
    segment = factor(segment)
  )
```

Remove variables that we used to compute segment. If we leave these variables in the data, the model will be able to perfectly predict `segment` - essentially, the model would be using `segment` to predict `segment`, which isn't useful.

```{r}
ifood <- ifood |>
  select(-c(AcceptedCmp3:AcceptedCmp2), -Response, -response_total)
```

Remove customer ID variable.

```{r}
ifood <- ifood |>
  select(-ID)
```

Replace 'nonsense' values of marital status with "Single".

```{r}
ifood |>
  distinct(Marital_Status)

nonsense_vals <- c("Alone", "Absurd", "YOLO")

ifood <- ifood |>
  mutate(Marital_Status = if_else(Marital_Status %in% nonsense_vals,
                                  "Single",
                                  Marital_Status))
```

## Pre-Processing

Create a new variable for the number of days since someone became a customer (i.e., customer age). As an example, let's pull out a single customer's date and compute how old they are as a customer.

```{r}
(date <- ifood |>
  slice(1) |>
  select(Dt_Customer))
```

Use `Sys.Date()` to reference today's date and compute the difference which is reported in days.

```{r}
date |>
  mutate(today = Sys.Date(),
         customer_age = as.numeric(today - Dt_Customer))
```

Repeat for the full data.

```{r}
ifood <- ifood |>
  mutate(customer_age = as.numeric(Sys.Date() - Dt_Customer)) |>
  select(-Dt_Customer)
```

Define train-test split and the cross-validation folds. Notice that the cross validation folds are created based on the training data only.

```{r}
set.seed(42)

ifood_split <- initial_split(ifood, prop = 0.75, strata = segment)

# Use v-fold cross-validation based on segment.
ifood_cv <- vfold_cv(training(ifood_split), v = 7, strata = segment)
```

Define models. We are going to have the baseline random forest and a tunable random forest. We'll compare their performance using cross validation and select the best model for evaluation on the test data.

```{r}
rf_model <- rand_forest() |> 
  set_engine(engine = "randomForest") |> 
  set_mode("classification")

rf_model_tuned <- rand_forest(mtry = tune(), 
                        trees = tune(), 
                        min_n = tune()) |> 
  set_engine(engine = "randomForest") |> 
  set_mode("classification")
```

Define recipe. We include a set for dummy variables for all nominal variables (this should apply to `Education` and `Marital Status`)

```{r}
# Feature engineering.
ifood_recipe <- training(ifood_split) |>
  recipe(segment ~ .) |>
  # create dummy variables for all non-numeric variables! (exclude the outcome - 
  # we don't want dummies for segment)
  # step_novel allows the recipe to account for factor levels (e.g., a dummy variable) that exists in the training data but not the validation or test data)
  step_dummy(all_nominal(), -segment)

ifood_recipe
```

Create two workflows, one for the baseline model and one for the tunable model.

```{r}
rf_workflow <- workflow() |>
  add_recipe(ifood_recipe) |>
  add_model(rf_model)

rf_workflow_tuned <- workflow() |>
  add_recipe(ifood_recipe) |>
  add_model(rf_model_tuned)
```

Train and evaluate the baseline model on the cross-validation folds.

```{r}
cv_results <- rf_workflow |> 
  fit_resamples(resamples = ifood_cv)
```

Do the same training and evaluation for the tunable model, but do it over a grid of hyperparameter values.

```{r}
cv_results_tuned <- rf_workflow_tuned |>
  tune_grid(resamples = ifood_cv)
```

Compute the average accuracy across folds for the baseline model.

```{r}
cv_results |>
  collect_metrics() |>
  filter(.metric == "accuracy")
```

We compute average accuracy across all folds for the tuned models and arrange them in descending order. The best model based on cross validation achieved 77% accuracy, compared to the 78% accuracy of the baseline model.

```{r}
cv_results_tuned |>
  collect_metrics() |>
  filter(.metric == "accuracy") |>
  arrange(desc(mean))
```

We make our modeling decision based on the cross-validation results. So, we select the baseline model as our final model, and evaluate it on the test data.

```{r}
# Fit the baseline workflow to the whole training dataset.
final_fit <- fit(rf_workflow, data = training(ifood_split))
```

Compute test data accuracy for the baseline model.

```{r}
# Compute model accuracy.
final_fit |> 
  predict(new_data = testing(ifood_split)) |>
  bind_cols(testing(ifood_split)) |>
  accuracy(truth = segment, estimate = .pred_class)
```

Our accuracy on this test data is consistent with the average accuracy across CV folds (within about 1%) so we can be confident that we are not overfitting the data.

Finally, we compute a confusion matrix.

```{r}
# Compute model accuracy.
final_fit |> 
  predict(new_data = testing(ifood_split)) |>
  bind_cols(testing(ifood_split)) |>
  conf_mat(truth = segment, estimate = .pred_class)
```

Let's break down each segment. For `Nonresponsive` customers, (second column), the model does a very good job at correctly classifying. It got 393 out of 403 correct. For `Semi-Responsive` and `Highly Responsive`, the model didn't do quite as well. The model had a little less than 50% accuracy for `Highly Responsive` customers, and very low accuracy for `Semi-Responsive` customers (only 11 out of 93). Ideally, we would want to find a model for iFood that wasn't as likely to classify `Highly Responsive` and `Semi-Responsive` as `Nonresponsive`, since these are valuable customers that the model is miss-classifying.
